Lifting CDCL to Relational Domains

### 2016-10-14 domains.tex
section 3.4

In case of partial assignments, meet irreducibles are the singleton
assignments, for example $\{x \mapsto true \}$, which represents the set 
of all propositional assignments where the literal $x$ is mapped to true.  
An important property of meet irreducibles in case of partial assignment is that 
they have precise complements.  For example, the complement of $\{x \mapsto true \}$
is a singleton element,$\{x \mapsto false \}$, in the partial assignment domain.

\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
complementable, then $A$ is said to be have complementable meet irreducibles.  
\end{definition}

%========================
\input{literals-clauses}
%========================

\subsection{Properties of Domain Elements for Decision and Learning}
%
An important property for a clause learning SAT solver 
is that each element of the partial assignments domain 
are decomposable into precisely complementable 
elements.  This property of the domain helps to 
guide the search away from the conflicting region of the 
search space.  
\rmcmt{define precise complement}
Thus, clause learning SAT solvers requires a weak complementation property
of the abstraction.  Most numerical abstract domain, such as intervals, 
octagons, polyhedra have complementation property, that is, they are
representable as intersections of complementable half-spaces. 
\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
Abstract Domain library, such as APRON C library~\cite{apron}, 
is specialized for the abstract interpretation framework which 
does not require complementation property.  To this end, we 
implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\Omit {
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise
operators.} } For example, the complementation of the octagon in Figure~\ref{octagon} 
can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

>>>>

A meet irreducible in partial assignments domain are the singleton 
assignments, for example $\{x \mapsto true \}$, which represents the set 
of all propositional assignments where the literal $x$ is mapped to true.  
A meet irreducible in interval domain is $\langle x \leq n \rangle$ or 
$\langle x \geq n \rangle$.

%
\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
complementable, then $A$ is said to be have complementable meet irreducibles.  
\end{definition}

\Omit {
An important property of meet irreducibles in case of partial assignments 
domain is that they have precise complements.  
For example, the complement of $\{x \mapsto true \}$ is a 
singleton element,$\{x \mapsto false \}$, in the partial assignments domain. 
}

%
%========================
\input{literals-clauses}
%========================

\subsection{Properties of Domain Elements for Learning}
%
\pscmt{TODO: Review by Peter}
An important property of a clause learning SAT solver 
is that each element of the partial assignments domain 
are decomposable into precisely complementable 
elements.  This property of the domain helps to guide 
the model search away from the conflicting region of the 
search space.

Most numerical abstract domains, such as intervals, octagons, 
polyhedra lacks precise complements, but they can be 
represented as intersections of complementable half-spaces, 
each of which have precise complements.  For example, intervals 
lack complements, but the decomposition of intervals into 
meet-irreducibles have complements.  
\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
Similarly, the complementation of the octagon in Figure~\ref{octagon} 
can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

Standard abstract interpretation does not require complementation property.  
Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
for abstract interpretation.  We implement a template-based polyhedra domain 
that returns precise complements by decomposing it into half-spaces.

\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
\Omit {
To this end, we implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise
operators.} 
}
